Convergence rate of MAP estimates \\ for the exponential family

### 2021-03-09 main.tex
No idea

\begin{figure}[h]
	\centering
	\includegraphics[width=.54\textwidth]{figs/asymptote.pdf}
	\includegraphics[width=.43\textwidth]{figs/fewsamples.pdf}
	\caption{Suboptimality of a gaussian variance MLE against number of samples $n$. Bold curve is average over 100 trials,  dark shaded area is 90\% (dark) confidence interval, ligth shade is min-max interval.  
	\textbf{Left:} as $n$ increases, the suboptimality matches the $1/2N$ asymptote.
	\textbf{Right:} the first few samples significantly deviate from this behavior, can we bound this deviation ? 
	}
	\label{fig:curves}
\end{figure}

\begin{example}
	\paragraph{Exact computation}
	The MLE of $\cN(0,\mu_*)$ is $\mu_n = \sum_i X_i^2 /n$. The ratio to the optimum  $\frac{\mu_n}{\mu_*}$follows a \href{https://en.wikipedia.org/wiki/Chi-square_distribution}{Chi-square distribution} with $n$ degrees of freedom $\chi^2(n)$ divided by $n$. Its inverse $\frac{\mu_*}{ \mu_n} $ follows an \href{https://en.wikipedia.org/wiki/Inverse-chi-squared_distribution}{inverse Chi-square distribution} with expectation 
	\begin{align}
		\expect{\frac{\mu_*}{ \mu_n} } =
		\begin{cases}
			\frac{n}{n-2} \ \text{ if } n>2, \\
			+\infty \  \text{ otherwise. }
		\end{cases}
	\end{align}
	There is also a closed form solution for the expected logarithm of a Chi-squared\citep{pav2015moments}
	\begin{align}
		\expect{\log \frac{\mu_n}{\mu_*}} = \log(2) + \psi(\half[n]) - \log(n)
	\end{align}
	where $\psi$ is the \href{https://en.wikipedia.org/wiki/Digamma_function}{digamma function.}
	Consequently the suboptimality of the MLE has a closed form solution
	\begin{align}
	\expect{\bregmanconj( \mu_*; \mu_n) }
	&= \half \expect{\frac{\mu_*}{ \mu_n} - 1 + \log(\frac{\mu_n}{\mu_*} } \\
	& =
	\begin{cases}
		\half \left ( \frac{2}{n-2} +\psi(\half[n]) - \log(\half[n]) \right )  \ \text{ if } n>2, \\
			+\infty \  \text{ otherwise. }
	\end{cases}
	\end{align}
	It is surprising that we need more than $3$ samples for the loss to have a bounded expectation. 
	It is also very pleasant that the optimal value $\mu_*$ does not appear in this rate, so that the convergence rate is independent of the actual solution.
	When the expectation is finite, we can make its formula intelligible thanks to \href{https://en.wikipedia.org/wiki/Digamma_function#Inequalities}{bounds on the digamma function}
	\begin{align}
		-\inv{x} \leq \psi(x) - \log(x) \leq -\inv{2x}	\; .
	\end{align}
	  For $n\geq 3$,
	\begin{align}
		\inv{n-2} - \inv{n}
		&\leq \expect{\bregmanconj( \mu_*; \mu_n) }
		\leq \inv{n-2} - \inv{2 n} \\
		\iff
			\frac{2}{n(n-2)}
			&\leq \expect{\bregmanconj( \mu_*; \mu_n) }
			\leq \inv{2n} +\frac{2}{n(n-2)}
			\label{eq:tight_bounds}
	\end{align}
	so we get a $O(n^{-2})$ lower bound and a $O(n^{-1}) + O(n^{-2})$ upper bound.
	Using the \href{https://en.wikipedia.org/wiki/Digamma_function#Recurrence_formula_and_characterization}{Stirling series}, we can also get an asymptotic approximation
	\begin{align}
		\expect{\bregmanconj( \mu_*; \mu_n) }
		\in \inv{2n} + \frac{11}{6 n^2} + \frac{4}{n^2(n-2)} + O(n^{-4}) \; .
	\end{align}
	Unfortunately, formulas for the MAP seem to be much less nice. There is also a closed form for the multivariate generalization, thanks to the \href{https://en.wikipedia.org/wiki/Inverse-Wishart_distribution}{inverse Wishart distribution} and the \href{https://en.wikipedia.org/wiki/Wishart_distribution#Log-expectation}{expectation of the log-determinant of a Wishart}. Again the expected value is infinite whenever $n \leq d+1$ where $d$ is the dimension. We do not perform these calculations for now as we want to focus on the simplest possible examples.
\end{example}

>>>>

\begin{figure}[h]
	\centering
	\includegraphics[width=.54\textwidth]{figs/asymptote.pdf}
	\includegraphics[width=.43\textwidth]{figs/fewsamples.pdf}
	\caption{Suboptimality of a Gaussian variance MLE against number of samples $n$. Bold curve is average over 100 trials,  dark shaded area is 90\% (dark) confidence interval, light shade is min-max interval.  
	\textbf{Left:} as $n$ increases, the suboptimality matches the $1/2N$ asymptote.
	\textbf{Right:} the first few samples significantly deviate from this behavior. In fact, for $n=1$ and $n=2$, the expected value is infinite, but we have a closed form solution and a simple upper-bound for $n>2$.
	}
	\label{fig:curves}
\end{figure}

\begin{example}
	\paragraph{Exact computation}
	The MLE of $\cN(0,\mu_*)$ is $\mu_n = \sum_i X_i^2 /n$. The ratio to the optimum  $\frac{\mu_n}{\mu_*}$follows a \href{https://en.wikipedia.org/wiki/Chi-square_distribution}{Chi-square distribution} with $n$ degrees of freedom $\chi^2(n)$ divided by $n$. Its inverse $\frac{\mu_*}{ \mu_n} $ follows an \href{https://en.wikipedia.org/wiki/Inverse-chi-squared_distribution}{inverse Chi-square distribution} with expectation 
	\begin{align}
		\expect{\frac{\mu_*}{ \mu_n} } =
		\begin{cases}
			\frac{n}{n-2} \ \text{ if } n>2, \\
			+\infty \  \text{ otherwise. }
		\end{cases}
	\end{align}
	There is also a closed form solution for the expected logarithm of a Chi-squared\citep{pav2015moments}
	\begin{align}
		\expect{\log \frac{\mu_n}{\mu_*}} = \log(2) + \psi(\half[n]) - \log(n)
	\end{align}
	where $\psi$ is the \href{https://en.wikipedia.org/wiki/Digamma_function}{digamma function.}
	Consequently the suboptimality of the MLE has a closed form solution
	\begin{align}
	\expect{\bregmanconj( \mu_*; \mu_n) }
	&= \half \expect{\frac{\mu_*}{ \mu_n} - 1 + \log \left(\frac{\mu_n}{\mu_*} \right) } \\
	& =
	\begin{cases}
		\half \left ( \frac{2}{n-2} +\psi(\half[n]) - \log(\half[n]) \right )  \ \text{ if } n>2, \\
			+\infty \  \text{ otherwise. }
	\end{cases}
	\end{align}
	It is surprising that we need more than $3$ samples for the loss to have a bounded expectation. 
	It is also very pleasant that the optimal value $\mu_*$ does not appear in this rate, so that the convergence rate is independent of the actual solution.
	When the expectation is finite, we can make its formula intelligible thanks to \href{https://en.wikipedia.org/wiki/Digamma_function#Inequalities}{bounds on the digamma function}
	\begin{align}
		-\inv{x} \leq \psi(x) - \log(x) \leq -\inv{2x}	\; .
	\end{align}
	  For $n\geq 3$,
	\begin{align}
		\inv{n-2} - \inv{n}
		&\leq \expect{\bregmanconj( \mu_*; \mu_n) }
		\leq \inv{n-2} - \inv{2 n} \\
		\iff
			\frac{2}{n(n-2)}
			&\leq \expect{\bregmanconj( \mu_*; \mu_n) }
			\leq \inv{2n} +\frac{2}{n(n-2)}
			\label{eq:tight_bounds}
	\end{align}
	so we get a $O(n^{-2})$ lower bound and a $O(n^{-1}) + O(n^{-2})$ upper bound.
	Using the \href{https://en.wikipedia.org/wiki/Digamma_function#Recurrence_formula_and_characterization}{Stirling series}, we can also get an asymptotic approximation
	\begin{align}
		\expect{\bregmanconj( \mu_*; \mu_n) }
		\in \inv{2n} + \frac{11}{6 n^2} + \frac{4}{n^2(n-2)} + O(n^{-4}) \; .
	\end{align}
	Unfortunately, formulas for the MAP seem to be much less nice. There is also a closed form for the multivariate generalization, thanks to the \href{https://en.wikipedia.org/wiki/Inverse-Wishart_distribution}{inverse Wishart distribution} and the \href{https://en.wikipedia.org/wiki/Wishart_distribution#Log-expectation}{expectation of the log-determinant of a Wishart}. Again the expected value is infinite whenever $n \leq d+1$ where $d$ is the dimension. We do not perform these calculations for now as we want to focus on the simplest possible examples.
\end{example}
